Untwisting the Tetrahelix

### 2017-04-11 doc/TetrahelixUntwist.tex
switch to SIAM style

\documentclass[11pt]{article}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{gensymb}
\usepackage[affil-it]{authblk}
\usepackage[toc,page]{appendix}
\usepackage{pifont}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{hyperref}

\usepackage{float}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

@@

\usepackage{draftwatermark}

\SetWatermarkText{DRAFT}
\SetWatermarkScale{6}
\SetWatermarkLightness{0.95}

@@

\title{Untwisting the Tetrahelix (v0.13)}
\author{Robert L. Read \texttt{read.robert@gmail.com}
}

@@

@@

The vertices of the tetrahedra lie upon
three helices about the central axis.  The
Tetrobot/Glussbot\cite{TetrobotBook} project uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  The Tetrobot concept is to use mechanical members, called
actuators, which can change their length, connected by special joints,
such as the three-d printable Song-Kwon-Kim\cite{song2003spherical} joint, or the CMS joint\cite{HamlinSandersonCMS}
which allow many members to come to a single point.
Such machines can
follow purely regular mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

@@

Imagining Figure \ref{fig:closeup} as a static mechanical structure,
we observe that it is useful to the mechanical engineer or
robotocist because the structure remains an inherently rigid,
omni-triangulated space frame, which is
somewhat mechanically strong.
Imagine further in Figure \ref{fig:closeup}, that each static edge was replaced with an
actuator that could dynamically become shorter or longer in response to electronic control,
and the vertices were a joint that supported sufficient angular displacement
for this to be possible. Such a machine is a Tetrobot\cite{TetrobotBook}.

A BC helix does not rest stably on a plane. It is convenient to
be able to ``untwist'' it and to form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, we can
untwist a tetrahelix to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile. This paper
develops the equations needed to untwist the tetrahelix. All math
developed here is available in JavaScript and demonstrated by an interactive
design website \url{https://pubinv.github.io/tetrahelix/}\cite{readtetrahelix},
from which Figure \ref{fig:closeup} and
the figures below are taken.

\begin{figure}
  \centering
     \includegraphics[width=1.0\textwidth]{figures/TetrahelixSeries2.png}
     \caption{A Series of Tetrahelices}
  \label{fig:series}
\end{figure}

Figure \ref{fig:series} displays a continuum of tetrahelices optimal in a certain sense, which is the result of this
paper. The closest helix is the BC helix, and the furthest
is the equitetrabeam, defined in Section \ref{sec:equitetrabeam}.

@@

We would like to design nearly regular tetrahelices with a formula that
gives the vertices in space. Eventually we would like to design nearly regular
tetrahelices by choosing the lengths of a small set of members.
In a space frame, this is a static design choice; in a robot, it is a
dynamic choice that can be used to twist\footnote{The formal definition of twistiness, or \emph{torsion},
is not useful or used in the paper. The formal \emph{curvature} and \emph{torsion} of the helices defined here
may be easily computed from the formualae if desired.} the robot and/or exert linear or
angular force on the environment.

@@

$\theta_{bc}$ is approximately $0.37 \cdot 2 \pi$ radians or  $131.81$ degrees.
The angle $\theta_{bc}$ is the rotation of \emph{each} tetrahedron, not the tetraheda along a rail.  In Figure \ref{fig:closeup},
each tetrahedron has either a yellow, blue, or red outer edge or rail.
That is, a blue-rail tetrahedron is rotated slightly more than a $1/3$ of a revolution to match the face of the yellow tetrahedra.

@@

Note in Figure \ref{railanglefig} the $z$-axis travel for one rail edge is denoted by $d$. In \eqref{graycoxeter} and \eqref{eq:colored}, the variable
$h$ is used for one third of the distance we name $d$. We will later justify that $d = 3h$.
In this paper we assume the length of a rail
is always $1$ as a simplification. (We make the rail length a parameter in our JavaScript code in \texttt{tetrahelixmath.js} \cite{readtetrahelix}.)

@@

Since all three rails have the same rail length, no matter how we
move the rails in the $xy$ plane if we shorten the $xy$ distance between
vertices we shorten the total distance.
Our tool for thinking about this is to collapse the $z$ dimension to form
a two-dimensional figure of nodes and non-rail edges (see Figure \ref{untouchingrailfig}.)
Consider the projection along the $z$ axis of all vertices and non-rail edges into the $xy$ plane, which will be
a figure of dots and connecting segments in the $xy$ plane. The convex
hull for any one helix projection will be a circle (if its pitch is
irrational) or a polygon if rational, or a point if the helix has
zero curvature. Each of these figures by definition lies outside the
circle of inradius $r_{in}$ in the $xy$ plane.

@@

Figure \ref{untouchingrailfig} depicts a projection of
only the point $P$, the point $P'$ which is an adjacent vertex on the same rail,
and the intermediary points $R$ and $Q$ on the other two rails.

Supoose there is a rail $P$ which does not have a segment touching the incircle
at all in the projection, as depicted in Figure \ref{untouchingrailfig}.
Then a segment connecting the other two $Q$ and $R$ is a longest
rail, because any chord touching an incircle is longer than any which does not.
These two rails ($Q$ and $R$) can be moved closer to each other, and further away from $P$, to form a better minimax solution
by shortening the longest rail, until one of the segments from $P$ the previously
untouching rail touches one of the incircle. By induction, then all rails have
at least one segement touching each vertex on the rail that is tangent to the
incircle in an optimal tetrahelix.

@@

Theorem \ref{thm:cylinder} is not true when $\rho$ is an odd integer multiple of $\pi$.

@@

Figure \ref{equitetrabeam} shows the equitetrabeam, which is defined in Section \ref{sec:equitetrabeam},
but illustrates the one-hop and two-hop edge defintions.
The green edges are the two-hop edges and the purple edges are the one-hop edges. Note that the green
edges are slightly longer than the purple edges.

@@

 Note that based on Theorem \ref{eventhirds}, there are only 3 possible lengths in an optimal tethrahelix,
 and we are justified in classifying edge lengths as \emph{rail}, \emph{one-hop}, or
\emph{two-hop}. The one-hop edges are the edges between rails that are closest on the $z$-axis, and the two-hop edges are those that skip over a vertex.

By Theorem \ref{eventhirds} every optimal tetrahelix has vertices lying on helices expressible in the form:
\[
V_{optimal}(n,c) =
\left [
  \begin{tabular}{c}
   $ r \cos(n \alpha +  c 2 \pi /3)$\\
   $ r \sin(n \alpha +  c 2 \pi /3) $\\
   $ \frac{d(n +c / 3)}{3}   $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $c \in \{0,1,2\}$
  \end{tabular}      
\]
where we have not yet investigated in the general case the relationships beteween $\alpha$, $r$, and $d$ in this formulation.
However, we understand that when $\alpha = 0$, the helices are degenerate, having curvature of $0$, and
we have the equitetrabeam.

@@

We seek a formula to generate optimal tetrahelices that accepts a
parameter that allows us to design the tetrahelix conveniently.
Please refer back to Figure \ref{railanglefig}.
The pitch of the helix is an obvious choice, but is not defined when the
curvature is $0$, an important special case. The radius or the axial
distance between two nodes on the same rail are possible choices, but
perhaps the clearest choice is to build formulae that takes as thier
input the ``rail angle'' $\rho$. We define $\rho$ to be the angle
formed in the X,Y plane $\angle H(0,0) O H(0,1)$ projecting out the $z$
axis and sighting along the positive $z$ axis. In other words, $\rho$
controls how far a rail edge of a tetrahelix deviates from being
parallel with the axis, or the ``twistiness'' of the tetrahelix. We use
the parameter $\chi = 1$ to indicate a chirality of counter-clockwise,
and $\chi = -1$ for clockwise.

@@

Rewriting our formulation in terms of $\rho$:
\begin{equation}
  \label{eq:general}
H_{general}(\chi,n,c,\rho,d_{\rho},r_{\rho}) =
\left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ r_{\rho}  \sin(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ d_{\rho} (n + c /3) $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $ \\
%%  $\kappa = n + c/3$ \\
    $\chi \in \{-1,1\}$ \\  
  \end{tabular}      
\end{equation}

@@

Note that when $\rho = 0$ then $h_{\rho} = 1$, but $r_{\rho}$ is not determined by
Equation \ref{railangle}.

@@

By Theoerm \ref{eventhirds}, we can compute the (at most) three edge-lengths of an optimal
tetrahelix by formula universally quantified by $n$ and $c$:
\begin{align*}
  \text{rail} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n+1,c,\rho,d_{\rho}),r_{\rho})) = 1 \\
  \text{one-hop} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n,c+1,\rho,d_{\rho},r_{\rho}))  \\
  \text{two-hop} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n,c+2,\rho,d_{\rho},r_{\rho}))  \\  
\end{align*}
where $dist$ is the Cartesian distance function.

@@

Let:
\begin{equation}
  f_{\rho} = -\frac{4 (\sin^2( \rho / 2))}{9}
  \end{equation}
\begin{equation}
  g_{\rho} = -\frac{16 (\sin^2( \rho / 2))}{9} 
\end{equation}

\begin{equation}
  j_{\rho} = \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)
\end{equation}
\begin{equation}
  k_{\rho} = (\sin^2(2\rho/3 + \frac{4\pi}{3})  + (1 - \cos(2\rho/3 + \frac{4\pi}{3}))^2)
\end{equation}

@@

\begin{align*}
  d_{opt}^2 &= 1 - 4 \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }}^2 (\sin{ \rho / 2})^2   \\
%%  d_{opt}^2 &= 1 - 4 \frac{4(\sin{ \rho / 2})^2}{\frac{9}{2} \cdot ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16(\sin{ \rho / 2})^2}{9 ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }    \\
%%  d_{opt}^2 &= 1 - \frac{16 \sin^2(\rho / 2) }{ \cos(\rho) + 9( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + 8 }   \\
    d_{opt} &= \sqrt{1 - \frac{16 \sin^2(\rho / 2) }{ \cos(\rho) + 9( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + 8 }}    \numberthis  \label{dopt}  \\      
\end{align*}

@@

If we look down the axis of an optimal tetrahelix as shown in Figure \ref{axialview}, it happens that only one of the one-hop edges
(rendered in purpple in our software)
comes closest to the axis. In other words, they define the radius of the incircle of the
projection, or the radius of a cylinder that would just fit inside the tetrahelix.
A formula for the inradious of the tetrahelix is useful if you are designing it as a structure that bears something internally,
such as a firehose, a pipe, or
a ladder for a human. The inradius $r_{in}(\rho)$ of
an optimal tetrahelix is a remarkably simple function of the radius $r$ and the rail angle $\rho$:
\begin{equation}
  \label{eq:inradius}
  r_{in}(\rho) = r \sin{\frac{\pi - \rho}{6}}
\end{equation}
Which can be seen from the trigonometry of a diagram of the projected one-hop edges
connecting four sequentially numbered vertices:

@@

Just as $H_{general}$ constructs the BC helix (with careful and non-obvious choices of parameters)
which is an important
special case due to its regularity, it constructs an additional special
(degenerate) case when the rail angle $\rho = 0$
and $d = 1$ (the edgelength), where the cross sectional area is
an equilateral triangle of unchanging orientation, as shown in Figure \ref{equitetrabeam} and at the rear of Figure \ref{fig:series}.
We call this the \emph{equitetrabeam}.

\begin{corollary}
  The equitetrabeam with minimal maximal edge difference is produced
  by $H_{general}$ when $ r = \sqrt{\frac{8}{27}} $.
\end{corollary}

@@

In Figure \ref{fig:series}, the furthest tetrahelix is the optimal equitetrabeam.
Note that folding the outer edges of the paper upward or downward effectively
chooses the chirality of the equitetrabeam.

@@

\begin{figure}[H] %float with two figures
  \centering
     \includegraphics[width=0.4\textwidth]{figures/GlussBotBC.jpg}
     \caption{Glussbot in relaxed, or BC helix configuration}
     \includegraphics[width=0.4\textwidth]{figures/GlussBotEquitetrabeamCropped.jpg}
     \caption{The Equitetrabream: Fully Untwisted Glussbot}
\end{figure}

@@

The Gluss Project \url{http://pubinv.github.io/gluss/} is part of Public Invention,
a free-libre, open-source research, hardware, and software project that welcomes volunteers.
It is our goal to organize projects for the benefit of all humanity without seeking profit or intellectual property.
To assist, contact \href{mailto:read.robert@gmail.com}{$<$read.robert@gmail.com$>$}.

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,gluss}

>>>>

\documentclass[review]{siamonline1116}
% \usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
% \usepackage{hyperref}
% \usepackage{graphicx}
\usepackage{gensymb}
% \usepackage[affil-it]{authblk}
% \usepackage[toc,page]{appendix}
% \usepackage{pifont}
\usepackage{amsmath}
% \usepackage{amsthm}
\usepackage{amsfonts}
% \usepackage{hyperref}

% \usepackage{float}

% \newtheorem{theorem}{Theorem}
% \newtheorem{corollary}{Corollary}

@@

% \usepackage{draftwatermark}

% \SetWatermarkText{DRAFT}
% \SetWatermarkScale{6}
% \SetWatermarkLightness{0.95}

@@

\title{Untwisting the Tetrahelix}
\author{Robert L. Read \email{read.robert@gmail.com}
}

@@

\begin{keywords}
  Boedijk--Coxeter helix, tetrahelix, robotics, tetrobot, unconventional robots,
  structural engineering, mechanical engineering, tensegrity, variable-geometry truss
\end{keywords}
\begin{AMS}
  51M15
\end{AMS}

@@

The vertices of the tetrahedra lie upon
three helices about the central axis.
The glussbot\cite{readglussbot} (or Tetrobot)\cite{TetrobotBook} uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  The Tetrobot concept is to use mechanical members, called
actuators, which can change their length, connected by special joints,
such as the three-d printable Song-Kwon-Kim\cite{song2003spherical} joint, or the CMS joint\cite{HamlinSandersonCMS}
which allow many members to come to a single point.
Such machines can
follow purely regular mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

@@

Imagining \cref{fig:closeup} as a static mechanical structure,
we observe that it is useful to the mechanical engineer or
robotocist because the structure remains an inherently rigid,
omni-triangulated space frame, which is
somewhat mechanically strong.
Imagine further in \cref{fig:closeup}, that each static edge was replaced with an
actuator that could dynamically become shorter or longer in response to electronic control,
and the vertices were a joint that supported sufficient angular displacement
for this to be possible. Such a machine is a Tetrobot\cite{TetrobotBook}.

A BC helix does not rest stably on a plane. It is convenient to
be able to ``untwist'' it and to form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, we can
untwist a tetrahelix to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile. This paper
develops the equations needed to untwist the tetrahelix. All math
developed here is available in JavaScript and demonstrated by an interactive
design website \url{https://pubinv.github.io/tetrahelix/}\cite{readtetrahelix},
from which \cref{fig:closeup} and
the figures below are taken.

\begin{figure}
  \centering
     \includegraphics[width=1.0\textwidth]{figures/TetrahelixSeries2.png}
     \caption{A Continuum of Tetrahelices}
  \label{fig:series}
\end{figure}

\Cref{fig:series} displays a continuum of tetrahelices optimal in a certain sense, which is the result of this
paper. The closest helix is the BC helix, and the furthest
is the equitetrabeam, defined in \cref{sec:equitetrabeam}.

@@

We would like to design nearly regular tetrahelices with a formula that
gives the vertices in space. Eventually we would like to design nearly regular
tetrahelices by choosing the lengths of a small set of members.
In a space frame, this is a static design choice; in a tetrobot, it is a
dynamic choice that can be used to twist\footnote{The formal definition of twistiness, or \emph{torsion},
is not useful or used in the paper. The formal \emph{curvature} and \emph{torsion} of the helices defined here
may be easily computed from the formualae if desired.} the robot and/or exert linear or
angular force on the environment.

@@

$\theta_{bc}$ is approximately $0.37 \cdot 2 \pi$ radians or  $131.81$ degrees.
The angle $\theta_{bc}$ is the rotation of \emph{each} tetrahedron, not the tetraheda along a rail.  In \cref{fig:closeup},
each tetrahedron has either a yellow, blue, or red outer edge or rail.
That is, a blue-rail tetrahedron is rotated slightly more than a $1/3$ of a revolution to match the face of the yellow tetrahedra.

@@

Note in \cref{railanglefig} the $z$-axis travel for one rail edge is denoted by $d$. In \eqref{graycoxeter} and \eqref{eq:colored}, the variable
$h$ is used for one third of the distance we name $d$. We will later justify that $d = 3h$.
In this paper we assume the length of a rail
is always $1$ as a simplification. (We make the rail length a parameter in our JavaScript code in \texttt{tetrahelixmath.js} \cite{readtetrahelix}.)

@@

Since all three rails have the same rail length, no matter how we
move the rails in the $xy$ plane if we shorten the $xy$ distance between
vertices we shorten the total distance.
Our tool for thinking about this is to collapse the $z$ dimension to form
a two-dimensional figure of nodes and non-rail edges (see \cref{untouchingrailfig}.)
Consider the projection along the $z$ axis of all vertices and non-rail edges into the $xy$ plane, which will be
a figure of dots and connecting segments in the $xy$ plane. The convex
hull for any one helix projection will be a circle (if its pitch is
irrational) or a polygon if rational, or a point if the helix has
zero curvature. Each of these figures by definition lies outside the
circle of inradius $r_{in}$ in the $xy$ plane.

@@

\Cref{untouchingrailfig} depicts a projection of
only the point $P$, the point $P'$ which is an adjacent vertex on the same rail,
and the intermediary points $R$ and $Q$ on the other two rails.

Supoose there is a rail $P$ which does not have a segment touching the incircle
at all in the projection, as depicted in \cref{untouchingrailfig}.
Then a segment connecting the other two $Q$ and $R$ is a longest
rail, because any chord touching an incircle is longer than any which does not.
These two rails ($Q$ and $R$) can be moved closer to each other, and further away from $P$, to form a better minimax solution
by shortening the longest rail, until one of the segments from $P$ the previously
untouching rail touches one of the incircle. By induction, then all rails have
at least one segement touching each vertex on the rail that is tangent to the
incircle in an optimal tetrahelix.

@@

\Cref{thm:cylinder} is not true when $\rho$ is an odd integer multiple of $\pi$.

@@

\Cref{equitetrabeam} shows the equitetrabeam, which is defined in \cref{sec:equitetrabeam},
but illustrates the one-hop and two-hop edge defintions.
The green edges are the two-hop edges and the purple edges are the one-hop edges. Note that the green
edges are slightly longer than the purple edges.

@@

 Note that based on \cref{eventhirds}, there are only 3 possible lengths in an optimal tethrahelix,
 and we are justified in classifying edge lengths as \emph{rail}, \emph{one-hop}, or
\emph{two-hop}. The one-hop edges are the edges between rails that are closest on the $z$-axis, and the two-hop edges are those that skip over a vertex.

By \cref{eventhirds} every optimal tetrahelix has vertices lying on helices expressible in the form:
\[
V_{optimal}(n,c) =
\left [
  \begin{tabular}{c}
   $ r \cos(n \alpha +  c 2 \pi /3)$\\
   $ r \sin(n \alpha +  c 2 \pi /3) $\\
   $ \frac{d(n +c / 3)}{3}   $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $c \in \{0,1,2\}$
  \end{tabular}      
\]
where we have not yet investigated in the general case the relationships beteween $\alpha$, $r$, and $d$ in this formulation.
However, we understand that when $\alpha = 0$, the helices are degenerate, having curvature of $0$, and
we have the equitetrabeam.

@@

We seek a formula to generate optimal tetrahelices that accepts a
parameter that allows us to design the tetrahelix conveniently.
Please refer back to \cref{railanglefig}.
The pitch of the helix is an obvious choice, but is not defined when the
curvature is $0$, an important special case. The radius or the axial
distance between two nodes on the same rail are possible choices, but
perhaps the clearest choice is to build formulae that takes as thier
input the ``rail angle'' $\rho$. We define $\rho$ to be the angle
formed in the X,Y plane $\angle H(0,0) O H(0,1)$ projecting out the $z$
axis and sighting along the positive $z$ axis. In other words, $\rho$
controls how far a rail edge of a tetrahelix deviates from being
parallel with the axis, or the ``twistiness'' of the tetrahelix. We use
the parameter $\chi = 1$ to indicate a chirality of counter-clockwise,
and $\chi = -1$ for clockwise.

@@

Rewriting our formulation in terms of $\rho$:
\begin{align*}
\numberthis  \label{eq:general}  
H_{general}(\chi,n,c,\rho,d_{\rho},r_{\rho}) &= 
 \left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ r_{\rho}  \sin(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ d_{\rho} (n + c /3) $
  \end{tabular}
  \right ] \\
\text{where: } &
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $ \\
    $\chi \in \{-1,1\}$ \\  
\end{tabular}
\end{align*}

@@

Note that when $\rho = 0$ then $h_{\rho} = 1$, but $r_{\rho}$ is not determined by
\cref{railangle}.

@@

By \cref{eventhirds}, we can compute the (at most) three edge-lengths of an optimal
tetrahelix by formula universally quantified by $n$ and $c$:
\begin{align*}
  \text{rail} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n+1,c,\rho,d_{\rho}),r_{\rho})) = 1 \\
  \text{one-hop} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n,c+1,\rho,d_{\rho},r_{\rho}))  \\
  \text{two-hop} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n,c+2,\rho,d_{\rho},r_{\rho}))  \\  
\end{align*}
where $dist$ is the Cartesian distance function.

@@

Let:
\begin{equation*}
  f_{\rho} = -\frac{4 (\sin^2( \rho / 2))}{9}
  \end{equation*}
\begin{equation*}
  g_{\rho} = -\frac{16 (\sin^2( \rho / 2))}{9} 
\end{equation*}

\begin{equation*}
  j_{\rho} = \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)
\end{equation*}
\begin{equation*}
  k_{\rho} = (\sin^2(2\rho/3 + \frac{4\pi}{3})  + (1 - \cos(2\rho/3 + \frac{4\pi}{3}))^2)
\end{equation*}

@@

\begin{align*}
  d_{opt}^2 &= 1 - 4 (\frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }})^2 (\sin{ \rho / 2})^2   \\
%%  d_{opt}^2 &= 1 - 4 \frac{4(\sin{ \rho / 2})^2}{\frac{9}{2} \cdot ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16(\sin{ \rho / 2})^2}{9 ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }    \\
%%  d_{opt}^2 &= 1 - \frac{16 \sin^2(\rho / 2) }{ \cos(\rho) + 9( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + 8 }   \\
    d_{opt} &= \sqrt{1 - \frac{16 \sin^2(\rho / 2) }{ \cos(\rho) + 9( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + 8 }}    \numberthis  \label{dopt}  \\      
\end{align*}

@@

If we look down the axis of an optimal tetrahelix as shown in \cref{axialview}, it happens that only one of the one-hop edges
(rendered in purpple in our software)
comes closest to the axis. In other words, they define the radius of the incircle of the
projection, or the radius of a cylinder that would just fit inside the tetrahelix.
A formula for the inradious of the tetrahelix is useful if you are designing it as a structure that bears something internally,
such as a firehose, a pipe, or
a ladder for a human. The inradius $r_{in}(\rho)$ of
an optimal tetrahelix is a remarkably simple function of the radius $r$ and the rail angle $\rho$:
\begin{equation}
  \label{eq:inradius}
  r_{in}(\rho) = r \sin{\frac{\pi - \rho}{6}}
\end{equation}
Which can be seen from the trigonometry of a diagram of the projected one-hop edges
connecting four sequentially numbered vertices:

@@

Just as $H_{general}$ constructs the BC helix (with careful and non-obvious choices of parameters)
which is an important
special case due to its regularity, it constructs an additional special
(degenerate) case when the rail angle $\rho = 0$
and $d = 1$ (the edgelength), where the cross sectional area is
an equilateral triangle of unchanging orientation, as shown in \cref{equitetrabeam} and at the rear of \cref{fig:series}.
We call this the \emph{equitetrabeam}.

\begin{corollary}
  The equitetrabeam with minimal maximal edge difference is produced
  \newline by $H_{general}$ when $ r = \sqrt{\frac{8}{27}} $.
\end{corollary}

@@

In \cref{fig:series}, the furthest tetrahelix is the optimal equitetrabeam.
Note that folding the outer edges of the paper upward or downward effectively
chooses the chirality of the equitetrabeam.

@@

\begin{figure}[H] %float with two figures
  \centering
     \includegraphics[width=0.9\textwidth]{figures/GlussBotBC.jpg}
     \caption{Glussbot in relaxed, or BC helix configuration}
     \includegraphics[width=0.9\textwidth]{figures/GlussBotEquitetrabeamCropped.jpg}
     \caption{The Equitetrabream: Fully Untwisted Glussbot}
\end{figure}

@@

The Gluss Project \url{http://pubinv.github.io/gluss/} is part of Public Invention \url{https://pubinv.github.io/PubInv/},
a free-libre, open-source research, hardware, and software project that welcomes volunteers.
It is our goal to organize projects for the benefit of all humanity without seeking profit or intellectual property.
To assist, contact \email{read.robert@gmail.com}.

\bibliographystyle{siamplain}
\bibliography{IEEEabrv,gluss}
